---
title: "Weather Station project"
output: html_document
date: "2024-03-14"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE,warning=FALSE)
```

```{r initialisation, include = FALSE}
library(here)
library(tidyverse)
library(gganimate)
#install.packages("gifski") # needed for gganimate
library(plotly) #roll over plots
library(htmlwidgets) #for saving
```


## Get the data!

Our data is here

https://www.metoffice.gov.uk/pub/data/weather/uk/climate/stationdata/sheffielddata.txt

Sheffield Weather Station in Western Park
```{r getdata}

source = here("data","sheffielddata.txt")

# this didn't work
#df <- read.csv(source,sep='\t',skip =7)

# we're going to load the file as text and clean it first
df <- readLines(source)
# clean some stuff
df <- (gsub("Provisional","",df))
df <- (gsub("\\*","",df))
df <- (gsub("\\#","",df))

# now save output as file
writeLines(df,here("data","modified.txt"))

#import from this
df <- read.table(here("data","modified.txt"),skip=7,fill=TRUE)

```

```{r wrangle}

cols <- c("year","month","max_t","min_t","af","rain","sun")

colnames(df) <-cols

df <- df %>%
  mutate(max_t = as.numeric(max_t))

#remember to come back and convert other vars to numeric

```

Now i am going to visualise some data

## Static plot

```{r plot1}

#create canvas
p <- ggplot(df,mapping = aes(x=month,y=max_t,group=year,col=year))
# add paint
p + geom_line(alpha=0.1)

```

## Animated plot!

This uses gganimate

<https://gganimate.com/>

```{r animation}

#new df which only contains average by year
cf <- df %>% group_by(year) %>%
  summarise(ave_upper = mean(max_t))

#canvas
p3 <- ggplot(cf,mapping = aes(x =year,y = ave_upper),frame = year)

#creat animation object
anim <- p3 + 
  geom_point() +
  transition_manual(year, cumulative = TRUE) +
  ggtitle('The year is: {frame + 1882}')
 
#show
anim 

#save plot
anim_save(here("plots","mygif.gif"),anim, renderer = gifski_renderer())
```

## ROLL OVER INTERACTIVE PLOT


This uses plotly

<https://plotly.com/ggplot2/>

```{r rollover}

#make plot object, the same as for the static plot
p <- ggplot(df,mapping = aes(x=month,y=max_t,group=year,col=year))
# note I make a new object which contains the 'canvas' + 'paint' together
p4 <- p + geom_line(alpha=0.1)


#show plot
ggplotly(p4)

#save
saveWidget(ggplotly(p4),file = here('plots','rollover.html'))

```

## Conclusion

It is nice to make animated plots, but they can distract as well as grab attention. Make sure the purpose of making the plot is served by adding animation.

## Colophon

Find the repo for this project here <https://github.com/tomstafford/station>

